### 2011-02-16 code/subarchitectures/dialogue.sa/branches/spring-school-2011/resources/grammars/openccg/prosodic_moloko.v6.0.1/docs/comsys.moloko/event-modifier-restriction.moloko.tex
branch for 2011 spring school

 \section{Event Modifier Restriction} \label{sec-Event-Modifier-Restriction}
We mentioned earlier that the Moloko is quite promiscous when it comes to  semantic or pragmatic 'un-acceptability', notions which must be handled grammar externally. There is, however, one major exception to this design principle: verbs are able to lexically specify what kinds of event modifiers they can recieve. This is the counterpart to treating most constituents as modifiers instead of as dependencies (see section X). In this section, first, we will give some examples illustrating the utility of this design feature, and then go into detail on how this was implemented.
\subsection{Motivation}
Consider first this pair of sentences:
\begin{itemize}
\item I played in the room
\item I walked in the room
\end{itemize}
Despite their similarity, these two sentences differ in that the second allows a reading which the first does not. 
In both cases, \emph{in the room} can be a static locational modifier, specifying the place where the event occured.\footnote{for the second, imagine a person pacing.} However, only in the second can \emph{in the room} specify the dynamic 'goal' of the event, i.e., walking \emph{into} the room. There is a case for arguing that this difference in motificational behaviour is inherit in the verb 
and should thus be lexically specified (c.f. Goldberg).  As another example, consider
\begin{itemize}
\item* I am a ball to the door
\end{itemize}
The unacceptability of this sentence can be accommodated by lexically specifying that the copular verb (be) blocks the class of 'dynamic modifiers'. Similarly, in the sentence
\begin{itemize}
\item* I want you to play in the room.
\end{itemize}
we can easily block the reading where \emph{in the room} modifies \emph{want}, i.e., the reading where it was the 'wanting' that occured within the room, instead of the true reading where it is the playing in the room that is wanted.

@@

The restriction of event modifiers is handled almost identically to the way the combinatorial behaviour of modifiers was handled (see \ref{sec-Dual-Relation-Words} above). It is done through the use of 
\begin{enumerate}
\item  a pair syntatic features which are attached lexically to modifiers and verbs
\item  the inclusion of these features within the syntactic category responsible for 'attaching' the  modifier to its modifee. This includes the type-changing rules used to handle  the modifier function of adverbs and prepositions (XXXXX), and also subordinate clause modifiers.\footnote{although event modifier restriction is surely a matter of semantics and not syntax, there are two reasons why it is better handled via feature. First, OpenCCG  much better at 'accessing' syntactic features than semantic ones. In fact, once semantic content has been added the semantic representation of a parse, it is no longer 'visible' to the parsing at all. Second, as discussed above (XXXXX), one of the design features for Moloko was to keep the semantic sorts as grammar external as possible. Using the event-type to control modification would violate this.}
\end{enumerate}
\subsubsection{Syntactic Features}
The first syntactic feature, m-type (modifier type), is attached to the event modifier and essentially parrots the modifier's semantic sort. For example, the preposition 'into', of semantic type 'm-whereto' is assigned a mod-type value of 's-whereto',

>>>>

 \section{Event Modifier Restriction}
As discussed in () above, the Moloko is quite promiscous when it comes to  semantic or pragmatic 'un-acceptability', notions which must be handled grammar externally. There is, however, one major exception to this design principle: verbs are able to lexically specify what kinds of event modifiers they can recieve. This is the counterpart to treating most constituents as modifiers instead of as dependencies (see section X). In this section, first, we will give some examples illustrating the utility of this design feature, and then go into detail on how this was implemented.
\subsection{Motivation}
Consider first this pair of sentences:
\begin{itemize}
\item I played in the room
\item I walked in the room
\end{itemize}
Despite their similarity, these two sentences differ in that the second allows a reading which the first does not. 
In both cases, \emph{in the room} can be a static locational modifier, specifying the place where the event occured.\footnote{for the second, imagine a person pacing.} However, only in the second can \emph{in the room} specify the dynamic 'goal' of the event, i.e., walking \emph{into} the room. There is a case for arguing that this difference in motificational behaviour is inherit in the verb 
and should thus be lexically specified (c.f. Goldberg).  As another example, consider
\begin{itemize}
\item* I am a ball to the door
\end{itemize}
The unacceptability of this sentence can be accommodated by lexically specifying that the copular verb (be) blocks the class of 'dynamic modifiers'. Similarly, in the sentence
\begin{itemize}
\item* I want you to play in the room.
\end{itemize}
we can easily block the reading where \emph{in the room} modifies \emph{want}, i.e., the reading where it was the 'wanting' that occured within the room, instead of the true reading where it is the playing in the room that is wanted.

@@

In fact, the restriction of event modifiers is handled almost identically to the way the combinatorial behaviour of modifiers was handled (see .. above). It is done through the use of 
\begin{enumerate}
\item  a pair syntatic features which are attached lexically to modifiers and verbs
\item  the inclusion of these features within the syntactic category responsible for 'attaching' the  modifier to its modifee. This includes the type-changing rules used to handle  the modifier function of adverbs and prepositions (), and also subordinate clause modifiers.\footnote{although event modifier restriction is surely a matter of semantics and not syntax, there are two reasons why it is better handled via feature. First, OpenCCG  much better at 'accessing' syntactic features than semantic ones. In fact, once semantic content has been added the semantic representation of a parse, it is no longer 'visible' to the parsing at all. Second, as discussed above (), one of the design features for Moloko was to keep the semantic sorts as grammar external as possible. Using the event-type to control modification would violate this.}
\end{enumerate}
\subsubsection{Syntactic Features}
The first syntactic feature, m-type (modifier type), is attached to the event modifier and essentially parrots the modifier's semantic sort. For example, the preposition 'into', of semantic type 'm-whereto' is assigned a mod-type value of 's-whereto',
